Viscous Erosion in Porous Media

### 2018-02-25 docs/methods/erosion_methods.tex
Editing paper

Our erosion model is based on the shear stress, and this model is
discussed in more detail in Section~\ref{sec:erosion}.  We then describe
how the incompressible Stokes equations are solved in $\Omega$ using a
boundary integral equation in Section~\ref{sec:bies}.  In
Section~\ref{sec:shearStressLP}, we discuss how the shear stress is
computed.  Finally, in Section~\ref{sec:thetaL}, we describe how corners
in the geometry are avoided and perform a change of coordinates to treat a stiff term.
\todo[inline]{I took out all references to the stiff term being linear, because with the new scaling we use (computing the mean shear stress on each body) the diffusive term is actually nonlinear (because the mean shear stress depends on theta). Nonetheless, the integrating-factor method still works.}

@@

Figure~\ref{shrink_intface}(a) shows that the spacing between successive interfaces becomes larger as the body shrinks, suggesting that the overall shear stress is growing. This observation is consistent with the well-known 3D Stokes drag law, $\tau \sim \mu \umax / L$, since stress is inversely proportional to body scale $L$. This law must be modified in two dimensions, though.
The modification is made non-trivial by subtleties of the 2D Stokes paradox, but can be correctly deduced by noting that, as the body vanishes, its drag must be consistent with slender-body theory \cite{poz1997, bouzarth2007epicyclic, camassa2008trajectory, MooreJFM2012, zhao2015transportation}. 
\todo[inline]{Here: add slender citations}
For a slender body in flow perpendicular to its long-axis $W \gg L$, the drag is proportional to
$\mu \umax W/ \log(W/L)$. This relationship implies that, for a 2D
cross-section, the stress per unit span can be estimated by
\begin{align}
  \label{eqn:stresslaw}
  \tau \sim \frac{\mu \umax}{L \log(W/L)}.
\end{align}
In interpreting our 2D simulations, we will treat the span, $W$, as fixed with the
requirement $W \gg L$.

The body's rate of area reduction is given by integrating the
absolute shear stress around the boundary
\begin{align*}
  \dot{A} = \int_{\gamma} \atau \, ds \, ,
\end{align*}
which gives the exact relationship
\begin{align*}
  \dot{A} = \mean{\atau} L \, .
\end{align*}
Thus, the area-loss rate is given by the product of the mean shear
$\mean{\atau}$ and the total perimeter $L$. Inserting the scaling
law~\eqref{eqn:stresslaw} gives
\begin{align*}
  \dot{A} \sim \frac{\mu \umax}{\log(W/L)} \, .
\end{align*}
Noting that $A \sim L^2$ allows one to solve the ODE to determine how the body area changes with time
\begin{align*}
  A (1 - \log A) = \areaconst (t_f - t)
\end{align*}
where $c_A$ is a constant.
This is an {\em implicit} relationship for how the area scales down in time during erosion.
In Fig.~6a, we show the body area as a function of time measured in the simulation
against the above scaling law with $\areaconst = 0.42$.

@@

We aim to measure the opening angle in our simulation, but, curiously,
the body vanishes before a steady-state is definitively reached. To
obtain the true terminal shape, we could restart our simulation with the
final eroded body scaled up and iterate this process until reaching
steady state. A simpler solution, though, is to remove the effect of
shrinking during the erosion process by simply modifying the normal
velocity $\Vn$ to have zero mean
\begin{align*}
  \Vn \leftarrow \Vn - \mean{\Vn}
\end{align*}
This modification allows erosion to modify the body shape as it would
normally, but artificially keeps the area of the body fixed. We show in
Fig.~\ref{fixed_intface}a several interfaces at equally spaced times.
Notice the body obtains the slender, football shape seen previously, but
here more precision is available. We also show in
Fig.~\ref{fixed_intface}b the stress, showing that the stress becomes
much more evenly distributed as time proceeds. In this figure, time is
normalized by the same vanishing time $t_f = 1.79E-2$ as the simulation
shown in Fig.~\ref{shrink_intface}. Curiously, about 3 multiples of
$t_f$ are required to reach a visually satisfying stead-state. This is
very much in contrast to previous studies of erosion in high-Reynold's
number flows, in which a reasonable steady-state was reached before the
body vanished~\cite{moo-ris-chi-zha-she2013}.

>>>>

Our erosion model is based on the shear stress, and this model is
discussed in more detail in Section~\ref{sec:erosion}.  We then describe
how the incompressible Stokes equations are solved in $\Omega$ using a
boundary integral equation in Section~\ref{sec:bies}.  In
Section~\ref{sec:shearStressLP}, we discuss how the shear stress is
computed.  Finally, in Section~\ref{sec:thetaL}, we describe how corners
in the geometry are avoided and perform a change of coordinates to treat a stiff term.

@@

Figure~\ref{shrink_intface}(a) shows that the spacing between successive interfaces becomes larger with time, suggesting that the overall shear stress is growing as the body vanishes. This observation is consistent with the well-known 3D Stokes drag law, $\tau \sim \mu \umax / L$, since stress is inversely proportional to body scale $L$. This law must be modified in two dimensions, though.
The modification is made non-trivial by subtleties of the 2D Stokes paradox, but can be correctly deduced by noting that, as the body vanishes, its drag must be consistent with slender-body theory \cite{poz1997, bouzarth2007epicyclic, camassa2008trajectory, MooreJFM2012, zhao2015transportation}. 
For a slender body in flow perpendicular to its long-axis $W \gg L$, the drag is proportional to
$\mu \umax W/ \log(W/L)$. This relationship implies that, for a 2D
cross-section, the stress per unit span can be estimated by
\begin{align}
  \label{eqn:stresslaw}
  \tau \sim \frac{\mu \umax}{L \log(W/L)}.
\end{align}
In interpreting our 2D simulations, we will treat the span, $W$, as fixed with the
requirement $W \gg L$.

The body's rate of area reduction is given by integrating the
absolute shear stress around the boundary
\begin{align*}
  \dot{A} = \int_{\gamma} \atau \, ds \,  = \mean{\atau} L \, .
\end{align*}
Thus, the area-loss rate is given by the product of the mean shear
$\mean{\atau}$ and the total perimeter $L$. Inserting scaling law~\eqref{eqn:stresslaw} gives
\begin{align*}
  \dot{A} \sim \frac{\mu \umax}{\log(W/L)} \, .
\end{align*}
Noting that $A \sim L^2$ allows one to solve the ODE to determine how the body area changes with time
\begin{align*}
  A (1 - \log A) = \areaconst (t_f - t)
\end{align*}
where $c_A$ is a constant.
This is an {\em implicit} relationship for how the area scales down in time during erosion.
In Fig.~6a, we show the body area as a function of time measured in the simulation
against the above scaling law with $\areaconst = 0.42$.

@@

We aim to measure the opening angle in our simulation, but, curiously, the body vanishes before a steady-state is definitively reached. To obtain the true terminal shape, we could restart our simulation with the final eroded body scaled up and iterate this process until reaching steady state. A simpler solution, though, is to remove the effect of shrinking during the erosion process by simply modifying the normal velocity $\Vn$ to have zero mean
\begin{align*}
  \Vn \leftarrow \Vn - \mean{\Vn}
\end{align*}
This modification allows erosion to modify the body shape as it would normally, but artificially keeps the area of the body fixed. We show in Fig.~\ref{fixed_intface}a several interfaces at equally spaced times. Notice the body obtains the slender, football shape seen previously, but here more precision is available. We also show in Fig.~\ref{fixed_intface}b the stress, showing that the stress becomes much more evenly distributed as time proceeds. In this figure, time is normalized by the same vanishing time $t_f = 1.79E-2$ as the simulation shown in Fig.~\ref{shrink_intface}. Curiously, about 3 multiples of $t_f$ are required to reach a visually satisfying stead-state. This is very much in contrast to previous studies of erosion in high-Reynold's number flows, in which an approximate steady-state was reached before the body vanished~\cite{moo-ris-chi-zha-she2013}.
